---
title: 4-Year Graduation Rate of Two High Schools
author: Foster, J.L.
date: 2019-03-25T00:00:00Z
summary: Visualization of the graduation rate of two similar high schools
tags:
- Visualization 
- School Outcomes
image:
  caption: Photo by  Jostens
  focal_point: Smart
---
I used public secondary data to visualize the difference in 4-year graduation rate for males and females attending two two similar high schools in the Pacific Northwest. Before creating these plots I combined five different data frames (years 2015-2019) of high school graduation in the entire state of Washington. After combining the data frames, I wrangled the data using dplyr from the tidyverse package in R. After filtering the dataframe to only capture my desired observations, I then used ggplot2 to create the graph, gganimate to animate the plots, and cowplot to make a multipanel plot of the two plots for ease of comparison.

```{r, message=FALSE}
#Packages I will use to graph the data
library(tidyverse)
library(ggrepel)
library(scales)
library(gganimate)
library(cowplot)
```

```{r, message=FALSE}
#Reading in Data
WAGrad15 <- read_csv("WAGrad14-15.csv")
WAGrad16 <- read_csv("WAGrad15-16.csv")
WAGrad17 <- read_csv("WAGrad16-17.csv")
WAGrad18 <- read_csv("WAGrad17-18.csv")
WAGrad19 <- read_csv("WAGrad18-19.csv")
```

```{r}
WAGrad <- bind_rows(WAGrad15, WAGrad16, WAGrad17, WAGrad18, WAGrad19)
```

```{r, include= FALSE}
SPS <-  filter(WAGrad, DistrictName == "Seattle School District #1", Cohort == 'Four Year')
```

```{r, include= FALSE}
SPS_Cut <- filter(SPS, !grepl('Non',StudentGroup))
SPS_Cut2 <- filter(SPS_Cut, !grepl('Students without',StudentGroup))
```

```{r, include= FALSE}
SPS_Cut3 <- select(SPS_Cut2,-c(ESDName,DistrictOrganizationId,CurrentSchoolType,Suppression,DataAsOf,TransferOut, Year1Dropout,Year2Dropout,Year3Dropout,Year4Dropout,Year5Dropout,Year6Dropout,Year7Dropout,SchoolOrganizationId,ESDOrganizationId,SchoolOrganizationid,ESDOrganizationID))
```

```{r, include= FALSE}
SPS_Cut4 <- filter(SPS_Cut3, OrganizationLevel == "School")
```

```{r}
#Using dplyr to wrangle the data to the exct data I want to graph
 ###I did several other filters witht this data
GBHS <- filter(SPS_Cut4, SchoolName %in% c("Ballard High School", "Garfield High School"), StudentGroupType %in% c("Race", "Gender", "All"))
```                 

```{r}

##Garfield and Ballard Overall Graduation Rate 
GBHS_Overall <- filter(GBHS, StudentGroupType %in% c("All", "Gender"))

##Ballard High School
BHS_Overall <- filter(GBHS,SchoolName %in% c("Ballard High School"),  StudentGroupType %in% c("Gender"))

##Garfield High School 
GHS_Overall <- filter(GBHS,SchoolName %in% c("Garfield High School"),  StudentGroupType %in% c("Gender"))

```

```{r}
##Animated Graduation rate for Ballard High School
ggplot(GHS_Overall, aes(x=SchoolYear, y=GraduationRate, color=StudentGroup)) + geom_line()  + theme(legend.position = c(.9,.9),legend.justification=  c("center","top"), legend.title= element_blank(),legend.key = element_blank(),legend.box.background = element_rect(color="black", size=.5), strip.background = element_rect(fill=NA, color=NA),panel.background = element_blank(),panel.grid.major = element_blank(), panel.grid.minor= element_blank(), panel.spacing.x=unit(.7, "lines"), panel.spacing.y=unit(.5, "lines")) + labs(x="\nGraduation Year", y= "Graduation Rate Percentage\n", title = "Garfield High School") + scale_y_continuous(labels= percent_format(accuracy = 1),limits = c(.80,1)) + scale_color_manual(values = c("red4", "navy")) + geom_point() +  transition_reveal(SchoolYear)

```

```{r}
##Animated Graduation rate for Ballard High School
ggplot(BHS_Overall, aes(x=SchoolYear, y=GraduationRate, color=StudentGroup)) + geom_line()  + theme(legend.position = c(.9,.9),legend.justification=  c("center","top"), legend.title= element_blank(),legend.key = element_blank(),legend.box.background = element_rect(color="black", size=.5), strip.background = element_rect(fill=NA, color=NA),panel.background = element_blank(),panel.grid.major = element_blank(), panel.grid.minor= element_blank(), panel.spacing.x=unit(.7, "lines"), panel.spacing.y=unit(.5, "lines")) + labs(x="\nGraduation Year", y= "Graduation Rate Percentage\n", title = "Ballard High School") + scale_y_continuous(labels= percent_format(accuracy = 1),limits = c(.80,1)) + scale_color_manual(values = c("red4", "navy")) + geom_point() + transition_reveal(SchoolYear)
```

```{r}
G <- ggplot(GHS_Overall, aes(x=SchoolYear, y=GraduationRate, color=StudentGroup)) + geom_line()  + theme(legend.position = c(0,.9),legend.justification=  c("center","top"), legend.title= element_blank(),legend.key = element_blank(),legend.box.background = element_rect(color="black", size=.5), strip.background = element_rect(fill=NA, color=NA),panel.background = element_blank(),panel.grid.major = element_line(color = "grey", size = 0.1), panel.grid.minor= element_blank(), panel.spacing.x=unit(.7, "lines"), panel.spacing.y=unit(.5, "lines")) + labs(x="\nGraduation Year", y= NULL, title = "Garfield High School") + scale_y_continuous(labels= percent_format(accuracy = 1),limits = c(.80,1)) + scale_color_manual(values = c("red4", "navy")) + geom_point() + theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

```

```{r}
B <- ggplot(BHS_Overall, aes(x=SchoolYear, y=GraduationRate, color=StudentGroup)) + geom_line()  + theme(legend.position = "none", legend.key = element_blank(),legend.box.background = element_rect(color="black", size=.5), strip.background = element_rect(fill=NA, color=NA),panel.background = element_blank(),panel.grid.major = element_line(color = "grey", size = 0.1), panel.grid.minor= element_blank(), panel.spacing.x=unit(.7, "lines"), panel.spacing.y=unit(.5, "lines")) + labs(x="\nGraduation Year", y= "Graduation Rate Percentage\n", title = "Ballard High School") + scale_y_continuous(labels= percent_format(accuracy = 1),limits = c(.80,1)) + scale_color_manual(values = c("red4", "navy")) + geom_point() + theme(panel.border = element_blank())
```

```{r}
#Multipanel Graph with Cowplot
plot_grid(B, G)
```

